---
title: "Dissertation - Data Analysis"
author: "Albatool"
date: "2/14/2023"
output:
  word_document: default
  html_document: default
---

```{r}
#install and load required packages
require(tidyverse)
require(readr)
library(dplyr)
library(tidytext)
require(psych)
require(lavaan)
require(semPlot)
library(kableExtra)
```

```{r}
#set seed to stablise bootstrap from computer clock 
#initial_seed = Sys.time()
#initial_seed=as.integer(initial_seed)
#print (initial_seed)
# take the last five digits of the initial seed
#the_seed=initial_seed %% 100000
#print(the_seed) #18144
set.seed(18144)

```


```{r}
#Reading the raw data file
#code line was deleted to preserve anonymity

```

# Data cleaning and preparing the data
```{r}
# data cleaning

#remove irrelevant columns 

clean_df <- numeric_df[,-c(1:6, 8:17, 47:50, 54:57, 61:64, 68:71, 75:78, 82:85, 89:92, 96:99, 103:106, 110:113, 117:120, 124:127, 131:134, 138:141, 145:148, 152:155, 159:162, 166:169, 173:176, 180:183, 187:190, 194:197, 201:204, 208:211, 215:218, 222:225, 229:232, 236:239, 243:246, 250:253, 278)]

#make first row the column names
colnames(clean_df) <- clean_df[1,] 
clean_df <- clean_df[-c(1,2), ]

#remove incomplete responses based on column "Finished" where 0=FALSE
summary(clean_df$Finished==0) #415 didn't complete study, 339 completed it
clean_df <- clean_df[!(clean_df$Finished==0),]
clean_df <- clean_df[,-1] #removing "Finished" columns after cleaning



#assigning unique ID number to each participant/row
clean_df <- dplyr::mutate(clean_df, ID = row_number())
clean_df <- clean_df %>% mutate(ID = factor(ID))
clean_df <- clean_df %>%
  relocate(ID)

```


```{r}
# data exclusions

#check if participants reported diagnosis with schizophrenia - if yes, exclude
clean_df[clean_df$`Have you been previously diagnosed with a neurological condition? - Yes (please specify if you feel comfortable sharing) - Text`%in% c('schizophrenia','Schizophrenia', 'Yes, schizophrenia')] 

clean_df[clean_df$`Have you been clinically diagnosed with a mental health condition? - Yes (please specify if you feel comfortable sharing) - Text`%in% c('schizophrenia','Schizophrenia', 'Yes, schizophrenia')] 

#also double-checked with manual search through df

  #no ppt reported diagnosis with schizophrenia. Through manual search found 1 ppt who reported schizoid diagnosis which will be excluded (ID #302)

#exclude schizoid case
clean_df <- clean_df[!(clean_df$ID==302),]

#acquired aphantasia 
clean_df[clean_df[,14] == '1',] #3 ppts indicated that their lack of visual mental imagery is acquired
summary(clean_df$`Was your difficulty with visual mental imagery (aphantasia) the result of a brain injury?`=='1') #just another way to check
clean_df <- clean_df[-c(30,124,191),] #removing ppts who answered yes by their unique id

  #NB: Q asked if aphantasia was acquired due to brain injury but aphantasia can be acquired with psychopathology; not very accurate formulation.

```


```{r}
# categorise aphantasia and phantasia groups based on VVIQ scoring

#sum score across rows for columns 15-30 and save score into new column (a score per ppt)
vviq <- clean_df %>% select(., c(15:30)) 
vviq <- sapply(vviq, as.numeric)
vviq <- as.data.frame(vviq)
vviq <- cbind(vviq, clean_df$ID)
vviq <- vviq %>% mutate(vviq_score = rowSums(across(.cols=c(1:16))))

#classify group based on scores
  #aphantasia = 75-80
  #phantasia = 24-74
  #hyperphantasia = 16-23

condition <- list()
score <- vviq$vviq_score

for(i in score) {
  if(i >=75){
    condition <- append(condition, 'aphantasia')
  } else if(i <= 23){
    condition <- append(condition, 'hyperphantasia')
  } else{
    condition <- append(condition, 'phantasia')
  }
}

#add group condition to dataset
condition<- as.factor(as.character(condition))
new_df <- clean_df %>% mutate(group = condition)
class(new_df$group)

#adding each ppts' vviq scores to df
new_df <- cbind(new_df, vviq$vviq_score)
new_df <- new_df %>% rename(vviq_score = `vviq$vviq_score`)

#check how many hyperphantasics
summary(new_df$group) #29 hyperphantasia

#exclude hyperphantasia condition
new_df<- new_df[new_df$group != 'hyperphantasia', ]
summary(new_df$group) #double-checking

#dummy code
new_df$condition <- ifelse(new_df$group == 'phantasia', 0, 1)
new_df <- new_df %>% mutate(condition=factor(condition))
levels(new_df$condition)
class(new_df$condition)




```



```{r}
# coding variables

#dummy code gender
new_df <- new_df %>% mutate(Gender=factor(Gender))
summary(new_df$Gender)

#if we were to exclude non-binary/third gender (14 ppts) and prefer not to say (3 ppts) - not enough data? - exclude and justify in write up but indicate original amount and how many were excluded

new_df <- new_df[new_df$Gender=='1' | new_df$Gender=='2', ] #keep only male and female
summary(new_df$Gender)

new_df$Gender <- ifelse(new_df$Gender=='1', 0, 1) #male as baseline
new_df$Gender <- as.factor(new_df$Gender)
class(new_df$Gender)

```

```{r}
first_language <- list()
lang <- new_df$`First language`

for(i in lang) {
  if(i =='English'){
    first_language <- append(first_language, '0')
  } else if(i=='english'){
    first_language <- append(first_language, '0')
  } else if(str_detect(i, 'english')){
    first_language <- append(first_language, '0')
  } else if(str_detect(i, 'English')){
    first_language <- append(first_language, '0')
  } else{
    first_language <- append(first_language, '1')
  }
}

first_language<- as.factor(as.character(first_language))
new_df <- new_df %>% mutate(first_language = first_language)
class(new_df$first_language)
summary(new_df$first_language)

sum(str_detect(new_df$`First language`, 'english')) #10
sum(str_detect(new_df$`First language`, 'English')) #205

#dummy code first language 
new_df$first_language <- ifelse(new_df$first_language=='0', 0,1)
new_df$first_language<- as.factor(new_df$first_language)
levels(new_df$first_language)

#missing 'enlish' and 'engish'

```

```{r}
#any transformations for years of education?
#exclude the 5 ppts who had ambiguous response? if lavaan accepts analysis with missing numerical values then keep

#dummy code diagnosis with neurological condition (1=yes, 2=no in the data - dummy code to 0=yes, 1=no)
new_df <- new_df %>% mutate(neurological_condition=factor(`Have you been previously diagnosed with a neurological condition? - Selected Choice`))

new_df$neuro_cond <- ifelse(new_df$neurological_condition=='1', 0,1)
new_df$neuro_cond <- as.factor(new_df$neuro_cond)

#dummy code diagnosis with mental health condition (1=yes, 2=no in data - dummy code to 0=yes, 1=no)
new_df <- new_df %>% mutate(mental_health_condition = factor(`Have you been clinically diagnosed with a mental health condition? - Selected Choice`))

new_df$mental_cond <- ifelse(new_df$mental_health_condition=='1', 0,1)
new_df$mental_cond <- as.factor(new_df$mental_cond)

```

```{r}
#reset unique ID number after deleting some ppts 
new_df <- new_df[,-1]
new_df <- dplyr::mutate(new_df, ID = row_number())
new_df <- new_df %>% mutate(ID = factor(ID))
new_df <- new_df %>%
  relocate(ID)

```

```{r}
#calculate anxiety score per ppt
#anxiety_items <- new_df %>% group_by(ID) %>% select(122,124,127,129,135,139,140) #gives errors when grouping
anxiety_items <- new_df %>% select(122,124,127,129,135,139,140)
anxiety_items <- as.data.frame(lapply(anxiety_items, as.numeric))
anxiety_items <- anxiety_items %>% mutate(anxiety_score=rowSums(.))

#multiply sum by 2 
anxiety_items <- anxiety_items*2

#calculate depression score per ppt 
depression_items <- new_df %>% select(123, 125, 130, 133, 136, 137, 141)
depression_items <- as.data.frame(lapply(depression_items, as.numeric))
depression_items <- depression_items %>% mutate(depression_score=rowSums(.))

#multiply sum by 2
depression_items <- depression_items*2

#add into df 
new_df <- cbind(new_df, anxiety_items$anxiety_score, depression_items$depression_score)
#seems like it's preserving order for each ppt

#need some descriptive stats for demographics 

```

```{r}
#calculate average scores per ppt across pictures for each emotional dimension 
#all SAM columns are class character, convert to class numeric 
new_df[31:120] <- sapply(new_df[31:120],as.numeric)

#valence scores
new_df$avg_valence <- rowMeans(new_df[ , c(31,34,37,40,43,46,49,52,55,58,61,64,67,70,73,76,79,82,85,88,91,94,97,100,103,106,109,112,115,118)], na.rm=TRUE)

#arousal scores
new_df$avg_arousal <- rowMeans(new_df[ , c(32,35,38,41,44,47,50,53,56,59,62,65,68,71,74,77,80,83,86,89,92,95,98,101,104,107,110,113,116,119)], na.rm=TRUE)

#dominance scores
new_df$avg_dominance <- rowMeans(new_df[ , c(33,36,39,42,45,48,51,54,57,60,63,66,69,72,75,78,81,84,87,90,93,96,99,102,105,108,111,114,117,120)], na.rm=TRUE)
```

```{r}
#remove redundant columns 
new_df <- new_df %>% select(c(1:14,142:154))

new_df <- new_df %>% rename(anxiety_score = `anxiety_items$anxiety_score`, depression_score = `depression_items$depression_score`, years_education = `Years of full-time education 
(include all years starting from primary school up until and including university)`)
```

```{r}
#make age numeric because it's character
new_df$Age <- as.numeric(new_df$Age)
class(new_df$Age)
#check variables are within possible range but that would be before removing the redundant columns

#convert years of education from character into numeric variable 
new_df$years_education <- as.numeric(new_df$years_education)
class(new_df$years_education)

summary(is.na(new_df$years_education)) #the model excluded those 5 participants with NA values
new_df <- na.omit(new_df)

```

# Descriptive Statistics 

```{r}
#descriptive stats of VVIQ scores 
new_df %>% group_by(group) %>% summarise(vviq_mean = round(mean(vviq_score),2),
                   vviq_sd = round(sd(vviq_score),2),
                   min_vviq = min(vviq_score),
                   max_vviq = max(vviq_score))

#distribution of vviq scores in all ppts
ggplot(new_df, aes(vviq_score)) +
  geom_histogram(aes(y = ..density..), fill="lightblue") +
  geom_density(alpha = 0.6, fill="grey") +
  labs(title="Distribution of VVIQ Scores",x="VVIQ Scores", y = "Density")+
  coord_cartesian(xlim = c(16, 80))

#distribution of vviq scores by group
new_df %>%
  ggplot( aes(x=vviq_score, fill=group)) +
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(fill="", title= substitute(paste(italic("Distribution of VVIQ Scores by Group"))),x="VVIQ Scores", y = "Count")+
  theme(text=element_text(size=17))+ 
  coord_cartesian(xlim = c(16, 80))


```



```{r}
new_df %>%
  summarise(mean_anxiety = round(mean(anxiety_score),2),
            mean_depression = round(mean(depression_score),2),
            mean_valence = round(mean(avg_valence),2),
            mean_arousal = round(mean(avg_arousal),2),
            mean_dominance = round(mean(avg_dominance),2)
            )
```


```{r}
#exploring descriptives
#contingency table of demographics summary 
new_df %>%
  group_by(group) %>%
  summarize(count=n(),
            `Mean Age` = round(mean(Age),2),
            #Gender = nrow(Gender),
            `Mean Education Years` = round(mean(years_education, na.rm=TRUE),2),
            #Education level,
            `First Language` = count(new_df,first_language),
            `Neurological Condition` = count(new_df,neuro_cond),
            `Mental Health Condition` = count(new_df,mental_cond)
            ) %>% t() %>% kable() %>% kable_classic() 

aggregate(cbind(Age, years_education) ~ group, data = new_df, mean)
aggregate(cbind(Gender, first_language, neuro_cond, mental_cond)~group, data=new_df, sum) 

#sjPlot::sjt.xtab(new_df$neuro_cond,new_df$group) 
gen_dem <- new_df %>% 
  group_by(group, Gender) %>%
  summarise(Gender = length(Gender))

mental_dem <- new_df %>% 
  group_by(group, mental_cond) %>%
  summarise(mental_health = length(mental_cond))

neuro_dem <- new_df %>% 
  group_by(group, neuro_cond) %>%
  summarise(neurological_condition = length(neuro_cond))

cbind(gen_dem,mental_dem, neuro_dem)

tb1 <- new_df %>% count(group,Gender)
tb2 <- new_df %>% count(group,neuro_cond)
cbind(tb1, tb2)


```

```{r}
#summary of neurological condition 
new_df <- new_df %>% mutate(neuro_diagnosis = as.factor(`Have you been previously diagnosed with a neurological condition? - Yes (please specify if you feel comfortable sharing) - Text`))
new_df %>% count(group, neuro_diagnosis) %>% kable() %>% kable_classic()

#summary of mental health conditions
new_df <- new_df %>% mutate(mental_health_diagnosis = as.character(`Have you been clinically diagnosed with a mental health condition? - Yes (please specify if you feel comfortable sharing) - Text`))
new_df %>% count(group, mental_health_diagnosis ) %>% kable() %>% kable_classic()

```


```{r}
#Testing normality of variables in the data  
new_df %>%
  group_by(condition) %>%
  summarise(statistic_age = shapiro.test(Age)$statistic,
            p.value_age = shapiro.test(Age)$p.value, 
            statistic_edu = shapiro.test(years_education)$statistic,
            p.value_edu = shapiro.test(years_education)$p.value
            ) 

#they are all significant and so violate normality - do non-parametric 
# two sample Wilcoxon test (also known as the Mann-Whitney test) and not independent t-test because not normally distributed 
#null=difference/distribution of two independent samples is the same; alternative=there is difference
MWW_age <- wilcox.test(Age~condition, data=new_df, paired=FALSE, alternative="two.sided")
#p-value<0.05 so we reject the null and conclude that the distribution of age between the two groups is different. The results indicate that [group] were significantly higher on age than [], W= , p<0.05. (if it's two sided do i need to see in which direction?)
MWW_edu <- wilcox.test(years_education~condition, data=new_df, paired=FALSE, alternative="two.sided")
#p-value=0.05 approaches significance but we fail to reject the null and conclude that the distribution of years of education is the not significantly different from zero/they are the same/there was no significant difference between aphantasics and phantasics on years of education.

#Mean, Median and SD of Age and Years of education
new_df %>% group_by(condition) %>% summarise(
  mean_age = round(mean(Age),2), 
  median_age = median(Age),
  sd_age = round(sd(Age),2),
  mean_edu.yrs = round(mean(years_education),2),
  median_edu.yrs = median(years_education),
  sd_edu.yrs = round(sd(years_education))
)

#checking mean age aphantasics reported realised their aphantasia
age_aphantasia <- as.numeric(as.character(new_df$`At what age did you realise having difficulties with visual mental imagery?`))
age_aphantasia <- age_aphantasia %>% na.omit()
age_aphantasia <- as.data.frame(age_aphantasia)
age_aphantasia %>% summarise(mean_age_aphantasia = round(mean(age_aphantasia),2))
```

#Fisher test interpretation:
Null (H0): There is no association between group and y. They are independent.
Alternative (H1): A relationship between group and y exists in the population.
When your p-value is below your significance level (e.g., 0.05), reject the null hypothesis. The sample data is strong enough to conclude that a relationship between the categorical variables exists in the population. Knowing the value of one variable provides information about the value of the other variable.
  
  
#Chi-squared test interpretation
a p-value that is less than or equal to your significance level indicates there is sufficient evidence to conclude that the observed distribution is not the same as the expected distribution. You can conclude that a relationship exists between the categorical variables.
```{r}
#Testing independence for nominal/categorical variables using Chi-square test 
#start by creating contingency table - gender~group
table(new_df$condition, new_df$Gender) %>% chisq.test()

#Neurological condition ~ group
table(new_df$condition, new_df$neuro_cond) %>% chisq.test() #non-significant p-value

#Mental health condition ~ group
table(new_df$condition, new_df$mental_cond) %>% chisq.test() #significant p-value

#First language ~ group 
table(new_df$condition, new_df$first_language) %>% chisq.test() #significant p-value


#highest education level ~ group 
  #factor first
new_df <- new_df %>% mutate(education_level = as.factor(`Highest level of qualified education you have completed`))
table(new_df$condition, new_df$education_level) %>% chisq.test() 
#significant p-value = there's overall difference
#post-hoc contrast analysis to determine where the effect is coming from 

levels(new_df$education_level)

#Level 1 vs. 2 = significant 
edu_1_vs_2 <- new_df[new_df$education_level %in% c("1", "2"), ] 
table(edu_1_vs_2$condition, edu_1_vs_2$education_level) %>% fisher.test() #chi-squared doesn't work; fisher's test

#Level 1 vs. 3 = non-significant 
edu_1_vs_3 <- new_df[new_df$education_level %in% c("1", "3"), ] 
table(edu_1_vs_3$condition, edu_1_vs_3$education_level) %>% fisher.test()

#Level 1 vs. 4 = non-significant 
edu_1_vs_4 <- new_df[new_df$education_level %in% c("1", "4"), ] 
table(edu_1_vs_4$condition, edu_1_vs_4$education_level) %>% fisher.test()

#Level 1 vs. 5 = non-significant
edu_1_vs_5 <- new_df[new_df$education_level %in% c("1", "5"), ] 
table(edu_1_vs_5$condition, edu_1_vs_5$education_level) %>% fisher.test()

#Level 1 vs. 6 = non-significant 
edu_1_vs_6 <- new_df[new_df$education_level %in% c("1", "6"), ] 
table(edu_1_vs_6$condition, edu_1_vs_6$education_level) %>% fisher.test()

#Level 2 vs. 3 = non-significant but approaches significance
edu_2_vs_3 <- new_df[new_df$education_level %in% c("2", "3"), ] 
table(edu_2_vs_3$condition, edu_2_vs_3$education_level) %>% fisher.test()

#Level 2 vs. 4 = significant 
edu_2_vs_4 <- new_df[new_df$education_level %in% c("2", "4"), ] 
table(edu_2_vs_4$condition, edu_2_vs_4$education_level) %>% fisher.test()

#Level 2 vs. 5 =  significant 
edu_2_vs_5 <- new_df[new_df$education_level %in% c("2", "5"), ] 
table(edu_2_vs_5$condition, edu_2_vs_5$education_level) %>% fisher.test()

#Level 2 vs. 6 = non-significant 
edu_2_vs_6 <- new_df[new_df$education_level %in% c("2", "6"), ] 
table(edu_2_vs_6$condition, edu_2_vs_6$education_level) %>% fisher.test()

#Level 3 vs. 4 = non-significant 
edu_3_vs_4 <- new_df[new_df$education_level %in% c("3", "4"), ] 
table(edu_3_vs_4$condition, edu_3_vs_4$education_level) %>% fisher.test()

#Level 3 vs. 5 = non-significant 
edu_3_vs_5 <- new_df[new_df$education_level %in% c("3", "5"), ] 
table(edu_3_vs_5$condition, edu_3_vs_5$education_level) %>% fisher.test()

#Level 3 vs. 6 = non-significant 
edu_3_vs_6 <- new_df[new_df$education_level %in% c("3", "6"), ] 
table(edu_3_vs_6$condition, edu_3_vs_6$education_level) %>% fisher.test()

#Level 4 vs. 5 = non-significant 
edu_4_vs_5 <- new_df[new_df$education_level %in% c("4", "5"), ] 
table(edu_4_vs_5$condition, edu_4_vs_5$education_level) %>% fisher.test()

#Level 4 vs. 6 = non-significant 
edu_4_vs_6 <- new_df[new_df$education_level %in% c("4", "6"), ] 
table(edu_4_vs_6$condition, edu_4_vs_6$education_level) %>% fisher.test()

#Level 5 vs. 6 - non-signifcant 
edu_5_vs_6 <- new_df[new_df$education_level %in% c("5", "6"), ] 
table(edu_5_vs_6$condition, edu_5_vs_6$education_level) %>% fisher.test()


#summary highest education level by group
table(new_df$condition, new_df$education_level)
```


```{r}
#Descriptive statistics for emotional processing by group 

#Testing normality of variables in the data before choosing test of independence 
new_df %>%
  group_by(group) %>%
  summarise(statistic_valence = round(shapiro.test(avg_valence)$statistic,2),
            p.value_valence = round(shapiro.test(avg_valence)$p.value,2),
            statistic_arousal = round(shapiro.test(avg_arousal)$statistic,2),
            p.value_arousal= round(shapiro.test(avg_arousal)$p.value,2),
            statistic_dominance = round(shapiro.test(avg_dominance)$statistic,2),
            p.value_dominance = round(shapiro.test(avg_dominance)$p.value,2)
            ) 

#they're all significant and so non-normal distribution except for phantasics' valence scores normal (p=0.8) but it's by group and at least one was significant (not normally distributed) 
  #Do non-parametric Mann-Whitney test 
  #null=difference/distribution of two independent samples is the same; alternative=there is difference
MWW_valence <- wilcox.test(avg_valence~group, data=new_df, paired=FALSE, alternative="two.sided") #no-sig
MWW_arousal <- wilcox.test(avg_arousal~group, data=new_df, paired=FALSE, alternative="two.sided") #sig
MWW_dominance <- wilcox.test(avg_dominance~group, data=new_df, paired=FALSE, alternative="two.sided") #non-sig

MWW_valence$statistic
round(MWW_valence$p.value,2)
MWW_arousal$statistic
MWW_arousal$p.value
MWW_dominance$statistic
round(MWW_dominance$p.value,2)

#Descriptives of emotional dimensions by group
new_df %>% group_by(group) %>% summarise(
  mean_valence = round(mean(avg_valence),2),
  sd_valence = round(sd(avg_valence),2),
  median_valence = round(median(avg_valence),2),
  min_valence = min(avg_valence),
  max_valence = max(avg_valence),
  mean_arousal = round(mean(avg_arousal),2),
  sd_arousal = round(sd(avg_arousal),2),
  median_arousal = round(median(avg_arousal),2),
  min_arousal = min(avg_arousal),
  max_arousal = max(avg_arousal),
  mean_dominance = round(mean(avg_dominance),2),
  sd_dominance = round(sd(avg_dominance),2),
  median_dominance = round(median(avg_dominance),2),
  min_dominance = min(avg_dominance),
  max_dominance = max(avg_dominance)
) 


```

```{r}
#Mental Health scores by group 

  #Testing normality of variables in the data before choosing test of independence 
new_df %>%
  group_by(group) %>%
  summarise(statistic_depression = round(shapiro.test(depression_score)$statistic,2),
            p.value_depression = round(shapiro.test(depression_score)$p.value,2),
            statistic_anxiety = round(shapiro.test(anxiety_score)$statistic,2),
            p.value_anxiety= round(shapiro.test(anxiety_score)$p.value,2)
            ) 

#they're all significant and so non-normal distribution 
  #Do non-parametric Mann-Whitney test 
  #null=difference/distribution of two independent samples is the same; alternative=there is difference
MWW_depression <- wilcox.test(depression_score~group, data=new_df, paired=FALSE, alternative="two.sided") #non-sig
MWW_anxiety <- wilcox.test(anxiety_score~group, data=new_df, paired=FALSE, alternative="two.sided") #non-sig

MWW_depression$statistic
MWW_depression$p.value
MWW_anxiety$statistic
MWW_anxiety$p.value

#Descriptive stats
new_df %>% group_by(group) %>% summarise(
  mean_depression = round(mean(depression_score),2),
  sd_depression = round(sd(depression_score),2),
  median_depression = round(median(depression_score),2),
  min_depression = min(depression_score),
  max_depression = max(depression_score),
  mean_anxiety = round(mean(anxiety_score),2),
  sd_anxiety = round(sd(anxiety_score),2),
  median_anxiety = round(median(anxiety_score),2),
  min_anxiety = min(anxiety_score),
  max_anxiety = max(anxiety_score)
) 


```


# Main Analyses
##Fitting path mediation models 

```{r}
#Model 1 - valence x depression 

mdl_1 <- '
avg_valence ~ a*condition + Age + Gender + first_language + years_education + neuro_cond 
depression_score ~ c*condition + b*avg_valence + Age + Gender + first_language + years_education + neuro_cond
ind:=a*b
dir:=c
total:= ind+dir
'

mdl_est_1 <- sem(mdl_1, data=new_df, se="bootstrap")
summary(mdl_est_1, ci=TRUE, fit.measures=T, standardized=TRUE)

#semPaths(mdl_est_1, what="what")

#Putting model results into table
standardizedSolution(mdl_est_1, ci=TRUE) %>% kable(caption="Path Mediation Model Results with the 95% Confidence Intervals",digits=2) %>% kable_paper()




```



```{r}
#model 2 - arousal x depression 
mdl_2 <- '
avg_arousal ~ a*condition + Age + Gender + first_language + years_education + neuro_cond 
depression_score ~ c*condition + b*avg_arousal + Age + Gender + first_language + years_education + neuro_cond 

ind:=a*b
dir:=c
total:= ind+dir
'

mdl_est_2 <- sem(mdl_2, data=new_df, se="bootstrap")
summary(mdl_est_2, ci=TRUE, fit.measures=T, std=TRUE)

semPaths(mdl_est_2, what="path")

standardizedSolution(mdl_est_2, ci=TRUE) %>% kable(caption="Path Mediation Model Results with the 95% Confidence Intervals",digits=2) %>% kable_paper()

```


```{r}
#model 3 - dominance x depression 
mdl_3 <- '
avg_dominance ~ a*condition + Age + Gender + first_language + years_education + neuro_cond 
depression_score ~ c*condition + b*avg_dominance + Age + Gender + first_language + years_education + neuro_cond 

ind:=a*b
dir:=c
total:= ind+dir
'

mdl_est_3 <- sem(mdl_3, data=new_df, se="bootstrap")
summary(mdl_est_3, ci=TRUE, fit.measures=T, std=TRUE)
standardizedSolution(mdl_est_3, ci=TRUE) %>% kable(caption="Path Mediation Model Results with the 95% Confidence Intervals",digits=2) %>% kable_paper()

#graph
#semPaths(mdl_est_3, what="path")
```


```{r}
#model 4 - valence x anxiety 
mdl_4 <- '
avg_valence ~ a*condition + Age + Gender + first_language + years_education + neuro_cond
anxiety_score ~ c*condition + b*avg_valence + Age + Gender + first_language + years_education + neuro_cond

ind:=a*b
dir:=c
total:= ind+dir
'

mdl_est_4 <- sem(mdl_4, data=new_df, se="bootstrap")
summary(mdl_est_4, ci=TRUE, fit.measures=T, std=TRUE)
standardizedSolution(mdl_est_4, ci=TRUE) %>% kable(caption="Path Mediation Model Results with the 95% Confidence Intervals",digits=2) %>% kable_paper()

#graph
#semPaths(mdl_est_4, what="path")
```

```{r}
#model 5 - arousal x anxiety 
mdl_5 <- '
avg_arousal ~ a*condition + Age + Gender + first_language + years_education + neuro_cond 
anxiety_score ~ c*condition + b*avg_arousal + Age + Gender + first_language + years_education + neuro_cond 

ind:=a*b
dir:=c
total:= ind+dir
'

mdl_est_5 <- sem(mdl_5, data=new_df, se="bootstrap")
summary(mdl_est_5, ci=TRUE, fit.measures=T, std=TRUE)
standardizedSolution(mdl_est_5, ci=TRUE) %>% kable(caption="Path Mediation Model Results with the 95% Confidence Intervals",digits=2) %>% kable_paper()

#graph
#semPaths(mdl_est_5, what="path")
```

```{r}
#model 6 - dominance x anxiety 
mdl_6 <- '
avg_dominance ~ a*condition + Age + Gender + first_language + years_education + neuro_cond  
anxiety_score ~ c*condition + b*avg_dominance + Age + Gender + first_language + years_education + neuro_cond 

ind:=a*b
dir:=c
total:= ind+dir
'

mdl_est_6 <- sem(mdl_6, data=new_df, se="bootstrap")
summary(mdl_est_6, ci=TRUE, fit.measures=T, std=TRUE)
standardizedSolution(mdl_est_6, ci=TRUE) %>% kable(caption="Path Mediation Model Results with the 95% Confidence Intervals",digits=2) %>% kable_paper()

#graph
#semPaths(mdl_est_6, what="path")
```


## Model Assumptions
```{r}
pairs.panels(new_df %>% select(Age, years_education, avg_valence, avg_arousal, avg_dominance, depression_score, anxiety_score))
#this is on the data rather than the model, bootstrap takes care of potential violation of linearity 


```

```{r}
#fitting separate regression models to test assumptions 
#lm(y~m+x) 
#lm(m~x) 
#model 1: 
mdl1_path1 <- lm(avg_valence~condition, data=new_df)
mdl1_path2 <- lm(depression_score~condition+avg_valence,data=new_df)

#model 2:
mdl2_path1 <- lm(avg_arousal~condition, data=new_df)
mdl2_path2 <- lm(depression_score~condition+avg_arousal, data=new_df)

#model 3:
mdl3_path1 <- lm(avg_dominance~condition, data=new_df)
mdl3_path2 <- lm(depression_score~condition+avg_dominance,data=new_df)

#model 4:
mdl4_path2 <- lm(anxiety_score~condition+avg_valence, data=new_df)

#model 5:
mdl5_path2 <- lm(anxiety_score~condition+avg_arousal, data=new_df)

#model 6:
mdl6_path2 <- lm(anxiety_score~condition+avg_dominance, data=new_df)
```

```{r}
#model 1 path 1
plot(mdl1_path1)

#model 1 path 2
plot(mdl1_path2) 



```

```{r}
new_df[new_df$ID==189,]
```

```{r}
#model 2 path 1
plot(mdl2_path1)

#model 2 path 2
plot(mdl2_path2)
```


```{r}
#model 3 path 1
plot(mdl3_path1)

#model 3 path 2
plot(mdl3_path2)
```

```{r}

#model 4 path 2
plot(mdl4_path2)
```

```{r}
#model 5 path 2
plot(mdl5_path2)

```

```{r}

#model 6 path 2
plot(mdl6_path2)

```


```{r}
#Plotting Mental Health Scores by Group
#Depression scores by group 
plt1 <- ggplot(data=new_df, aes(x = group, y = depression_score, fill=group)) +
    stat_boxplot(geom = "errorbar", width=0.1) + 
  geom_boxplot(outlier.colour = "#566573",outlier.size = 2, alpha=.3, fatten=.25)+
  geom_point(data=new_df, aes(y=depression_score, colour=group), alpha=0.3, position=position_jitterdodge(), show.legend=FALSE)+
  stat_summary(fun="mean", shape=1, size=0.5, show.legend=FALSE) +
  labs(title="Depression Scores by Group", x="Group", y="Depression Scores", fill="")+
  scale_x_discrete(limits=c("aphantasia","phantasia"),labels=c("aphantasia"="Aphantasia","phantasia"="Phantasia"))+
  theme_classic()+theme(text=element_text(size=15))+
  theme(legend.position="bottom")

#Anxiety scores by group 
plt2<- ggplot(data=new_df, aes(x = group, y = anxiety_score, fill=group)) +
    stat_boxplot(geom = "errorbar", width=0.1) + 
  geom_boxplot(outlier.colour = "#566573",outlier.size = 2, alpha=.3, fatten=.25)+
  geom_point(data=new_df, aes(y=anxiety_score, colour=group), alpha=0.3, position=position_jitterdodge(), show.legend=FALSE)+
  stat_summary(fun="mean", shape=1, size=0.5, show.legend=FALSE) +
  labs(title="Anxiety Scores by Group", x="Group", y="Anxiety Scores", fill="")+
  scale_x_discrete(limits=c("aphantasia","phantasia"),labels=c("aphantasia"="Aphantasia","phantasia"="Phantasia"))+
  theme_classic()+theme(text=element_text(size=15))+
  theme(legend.position="bottom")

cowplot::plot_grid(plt1, plt2, labels = c('A', 'B'), label_size = 12)

```


```{r}
#Plotting Emotional Dimensions by Group
#Used group variable here but it's the same with condition variable (dummy coded version)
#although group variable has hyperphantasia as third level, because all data-points were removed it doesn't matter which variable we use, mean points and distribution is the same

#Average valence scores by group 
p1<- ggplot(data=new_df, aes(x = group, y = avg_valence, fill=group)) +
    stat_boxplot(geom = "errorbar", width=0.1) + 
  geom_boxplot(outlier.colour = "#566573",outlier.size = 2, alpha=.3, fatten=.25)+
  geom_point(data=new_df, aes(y=avg_valence, colour=group), alpha=0.3, position=position_jitterdodge(), show.legend=FALSE)+
  stat_summary(fun="mean", shape=1, size=0.5, show.legend=FALSE) +
  labs(title="Mean Valence Scores by Group", x="Group", y="Valence Score", fill="")+
  scale_x_discrete(limits=c("aphantasia","phantasia"),labels=c("aphantasia"="Aphantasia","phantasia"="Phantasia"))+
  theme_classic()+theme(text=element_text(size=15))+
  theme(legend.position="bottom")


#Average arousal scores by group 
p2<- ggplot(data=new_df, aes(x = group, y = avg_arousal, fill=group)) +
    stat_boxplot(geom = "errorbar", width=0.1) + 
  geom_boxplot(outlier.colour = "#566573",outlier.size = 2, alpha=.3, fatten=.25)+
  geom_point(data=new_df, aes(y=avg_arousal, colour=group), alpha=0.3, position=position_jitterdodge(), show.legend=FALSE)+
  stat_summary(fun="mean", shape=1, size=0.5, show.legend=FALSE) +
  labs(title="Mean Arousal Scores by Group", x="Group", y="Arousal Score", fill="")+
  scale_x_discrete(limits=c("aphantasia","phantasia"),labels=c("aphantasia"="Aphantasia","phantasia"="Phantasia"))+
  theme_classic()+theme(text=element_text(size=15)) +
  theme(legend.position="bottom")


#Average dominance scores by group 
p3<- ggplot(data=new_df, aes(x = group, y = avg_dominance, fill=group)) +
    stat_boxplot(geom = "errorbar", width=0.1) + 
  geom_boxplot(outlier.colour = "#566573",outlier.size = 2, alpha=.3, fatten=.25)+
  geom_point(data=new_df, aes(y=avg_dominance, colour=group), alpha=0.3, position=position_jitterdodge(), show.legend=FALSE)+
  stat_summary(fun="mean", shape=1, size=0.5, show.legend=FALSE) +
  labs(title="Mean Dominance Scores by Group", x="Group", y="Dominance Score", fill="")+
  scale_x_discrete(limits=c("aphantasia","phantasia"),labels=c("aphantasia"="Aphantasia","phantasia"="Phantasia"))+
  theme_classic()+theme(text=element_text(size=15))+ 
  theme(legend.position="bottom")

cowplot::plot_grid(p1, p2, p3, labels = c('A', 'B', 'C'), label_size = 12)
```




